---
title: "Early Childhood"
---

```{r}
#
source(here::here("setup.R"))
source(here::here("footnotes.R"))
```

::: {.caveat-box}
**A note about the impact of under-identification**

`r note_box_underidentification`
:::

:::{style="height: 2em;"}
:::

::: {.key-question}
Are Aboriginal and/or Torres Strait Islander children living in the NEPHU catchment getting the best start in life?
:::

# Preschool

::: {.ctg-box}
**This topic aligns with [Closing the Gap Outcome Area 3](https://www.pc.gov.au/closing-the-gap-data/dashboard/outcome-area/early-childhood-education/): Children are engaged in high quality, culturally appropriate early childhood education in their early years**

Target: By 2025, increase the proportion of Aboriginal and Torres Strait Islander children enrolled in Year Before Full-Time Schooling (YBFS) early childhood education to 95%.
:::

:::{style="height: 1em;"}
:::

```{r}
# ------------------------------------------------------------------------------
# Preschool enrollments
# Data from Victorian Child and Adolescent Monitoring System
# ------------------------------------------------------------------------------
# Read in VCAMS data
childhood_enrol <- readxl::read_xlsx(file.path(root_folder, subfolder_vcams, "VCAMS_indicator_4_3c.xlsx"), 
                                     sheet        = "4.3c 2018 onwards", 
                                     trim_ws      = TRUE, 
                                     .name_repair = "universal_quiet") %>% 
  #
  janitor::clean_names()

# Data wrangling
childhood_enrol <- childhood_enrol %>%
  dplyr::select(year,
                aboriginal_n = numerator,
                aboriginal_prop = indicator_calc) %>% 
  #
  dplyr::mutate(aboriginal_prop = aboriginal_prop * 100)

# ------------------------------------------------------------------------------
# Preschool attendance
# Data from National Early Childhood Education and Care Collection
# ------------------------------------------------------------------------------
# Read in NECECC data
childhood_attendance <- readxl::read_excel(file.path(root_folder, subfolder_ctg,
                                                     "CTG03_Childhood_Education_Data_Tables.xlsx"),
                                           skip  = 1,
                                           sheet = "Table SE3a.1") %>% 
    janitor::clean_names() %>% 
    #
    dplyr::rename(value = x3,
                  indigenous_status = x4,
                  year = x5,
                  attendance = x6) %>%
    #
    tidyr::fill(value, indigenous_status, year, attendance) %>% 
    #
    dplyr::filter(!is.na(vic)) %>%
    #
    dplyr::select(value,
                  year,
                  indigenous_status,
                  attendance,
                  vic)

# Data wrangling
childhood_attendance <- childhood_attendance %>% 
  dplyr::mutate(vic = as.numeric(vic),
                #
                attendance = factor(dplyr::case_when(
                  attendance == "Did not attend"            ~ "Did not attend",
                  attendance == "One to less than 10 hours" ~ "<10 hours",
                  attendance == "10–14 hours"               ~ "10-14 hours",
                  attendance == "15 hours or more"          ~ "15+ hours",
                  #
                  TRUE ~ NA_character_),
                  #
                  levels = c("Did not attend",
                             "<10 hours",
                             "10-14 hours",
                             "15+ hours")))
```

- Data for this topic were obtained from the [Victorian Child and Adolescent Monitoring System (VCAMS)](https://www.vic.gov.au/victorian-child-and-adolescent-monitoring-system-vcams) and the [Closing the Gap Information Repository](https://www.pc.gov.au/closing-the-gap-data/).

- In 2022, 100% of Aboriginal and/or Torres Strait Islander children in Victoria who were in the year before full-time schooling age cohort attended a funded kindergarten/preschool program. This is above the Closing the Gap target of 95%.

- The percentage of enrolled children who attended for 15 or more hours in a week increased from 64% in 2016 to 75% in 2023.

**Key points from stakeholder consultation**

- At stakeholder consultations during the development of this report the impact of the scaling up of free kindergarten to 15 hours per week in 2022 was discussed as a potential contributing factor to rising preschool attendance.

- The importance of and impact of improvements in cultural safety by educators was also discussed as a potential contributing factor to this increase.

:::: {.panel-tabset}




::::

