---
title: "Diabetes"
---

```{r}
#
source(here::here("setup.R"))
source(here::here("setup_footnotes.R"))

# Set up numbering for exporting tables as .csv
table_section <- "10F"
table_count   <- 0

figure_count <- 0
```

::: {.caveat-box}
**A note about the impact of under-identification**

`r note_box_underidentification`

**A note about interpreting observed vs. expected figures**

`r note_box_obsexp`
:::

<div style="height: 2em;"></div>

::: {.key-question}
What chronic health conditions are experienced by Aboriginal and/or Torres Strait Islander peoples in the NEPHU catchment?
:::

## Diabetes

### Self-Reported Diabetes

```{r}
# ------------------------------------------------------------------------------
# Diabetes, by LGA
# Data from ABS Census using TableBuilderPro
# ------------------------------------------------------------------------------
# Read in Census data
diabetes_lga <- f_read_census_lga("Aboriginal_Health_Diabetes_LGA.csv",
                                  variable = "diabetes") %>% 
  #
  dplyr::filter(diabetes != "Not stated")

# Data wrangling
diabetes_lga <- diabetes_lga %>%
  #
  dplyr::mutate(diabetes = dplyr::case_when(
    diabetes == "Has diabetes (excluding gestational diabetes)" ~ "yes",
    TRUE ~ "no")) %>% 
  #
  f_wrangle_census_lga(variable = diabetes,
                       level    = "person",
                       dichot   = "yes")

# ------------------------------------------------------------------------------
# Diabetes, by age and sex
# Data from ABS Census using TableBuilderPro
# ------------------------------------------------------------------------------
# Read in Census data
diabetes_agesex <- f_read_census_agesex("Aboriginal_Health_Diabetes_AgeSex.csv",
                                        variable = "diabetes") %>% 
  #
  dplyr::filter(diabetes != "Not stated")

# Data wrangling
diabetes_agesex <- diabetes_agesex %>% 
  #
  dplyr::mutate(diabetes = dplyr::case_when(
    diabetes == "Has diabetes (excluding gestational diabetes)" ~ "yes",
    TRUE ~ "no")) %>% 
  #
  f_wrangle_census_agesex(variable = diabetes,
                          dichot   = "yes")
```

- Data for this topic were obtained from the [ABS Census 2021](https://www.abs.gov.au/census).

- Overall, 5.0% (n = 566) of people living in the NEPHU catchment who identified as Aboriginal and/or Torres Strait Islander in the Census have been told by a doctor or nurse that they have diabetes. This was similar to non-Indigenous people living in the NEPHU catchment (4.8%).

- The number of people living in different LGAs within the NEPHU catchment who identified as Aboriginal and/or Torres Strait Islander in the Census who have been told by a doctor or nurse that they have diabetes ranged from 97 (Hume LGA) to 5 (Manningham LGA).

- The gap between the observed number of Aboriginal and/or Torres Strait Islander people who have been told by a doctor or nurse that they have diabetes and the number expected if the rate was the same as for non-Indigenous people was biggest for Whittlesea and Darebin LGAs.

- The percentage of people living in different LGAs within the NEPHU catchment who identified as Aboriginal and/or Torres Strait Islander in the Census who have been told by a doctor or nurse that they have diabetes ranged from 6.9% (Darebin LGA) to 1.8% (Manningham LGA).

- The number of people who have been told by a doctor or nurse that they have diabetes was highest for people aged 50-59 or 60-69 years for both males and females, and the percentage of people who have been told by a doctor or nurse that they have diabetes increased with increasing age.

- The observed number of Aboriginal and/or Torres Strait Islander people who have been told by a doctor or nurse that they have diabetes was particularly higher than the number expected if the rate was the same as for non-Indigenous people for all age groups less than 70 years.

**Key points from stakeholder consultation**

- At stakeholder consultation during the development of this report, the association between healthy eating and the risk of diabetes was discussed.

- It was also noted that gestational diabetes is a separate condition that is not included in these data.

::: {.callout-tip title="Question/s asked for this topic" icon=false collapse=false}
The following question was asked of all people completing the ABS Census 2021.

- Has the person been told by a doctor or nurse that they have any of these long-term health conditions? (Select all that apply)
  - Arthritis
  - Asthma
  - Cancer (including remission)
  - Dementia (including Alzheimer's)
  - Diabetes (excluding gestational diabetes)
  - Heart disease (including heart attack or angina)
  - Kidney disease
  - Lung condition (including COPD or emphysema)
  - Mental health condition (including depression or anxiety)
  - Stroke
  - Any other long-term health condition(s)
  - No long-term health condition

<div style="height: 2em;"></div>
:::

:::: {.panel-tabset}

#### Figure 10F-1

::: {#fig_10F_1 .fig-tab-title}
**Figure 10F-1:** Number and percentage of people who identified as Aboriginal and/or Torres Strait Islander and who have been told by a doctor or nurse that they have diabetes, NEPHU catchment, by LGA, 2021
:::

```{r}
#
figure_count <- figure_count + 1

figure_number <- diabetes_lga %>% 
  f_bar_simple_lga(y_variable = aboriginal_n,
                   y_breaks   = 10,
                   y_expand   = 5)

figure_percent <- diabetes_lga %>% 
  f_bar_simple_lga(y_variable = aboriginal_prop,
                   y_max      = 10,
                   y_breaks   = 1,
                   y_expand   = 0.25)

f_bar_simple_combine(figure_number, figure_percent)
```

#### Figure 10F-2

::: {#fig_10F_2 .fig-tab-title}
**Figure 10F-2:** Observed vs. expected number of Aboriginal and/or Torres Strait Islander people who have been told by a doctor or nurse that they have diabetes, NEPHU catchment, by LGA, 2021
:::

```{r}
#
figure_count <- figure_count + 1

diabetes_lga %>% 
  f_bar_obsexp_lga(y_breaks = 25,
                   y_expand = 8)
```

#### Table 10F-1

::: {#tab_10F_1 .fig-tab-title}
**Table 10F-1:** Observed vs. expected number of Aboriginal and/or Torres Strait Islander people who have been told by a doctor or nurse that they have diabetes, NEPHU catchment, by LGA, 2021
:::

```{r}
#
table_count  <- table_count + 1
table_number <- paste0("Table_", table_section, "_", sprintf("%.02d", table_count))

table <- diabetes_lga %>%
  f_preptable_obsexp_lga()

if (!file.exists(paste0(here::here(), "/data/", table_number, ".csv"))) {
  
  write.csv(table, paste0(here::here(), "/data/", table_number, ".csv"), row.names = FALSE)
  
}

table %>% 
  f_table_obsexp(column_name = "LGA",
                 header_name = " ")
```

<div style="height: 1em;"></div>

<a href="../data/`r paste0(table_number, '.csv')`" download class="btn btn-primary download-link">Download table data as CSV</a>

#### Figure 10F-3

::: {#fig_10F_3 .fig-tab-title}
**Figure 10F-3:** Number and percentage of people who identified as Aboriginal and/or Torres Strait Islander and who have been told by a doctor or nurse that they have diabetes, NEPHU catchment, by age and sex, 2021
:::

```{r}
#
figure_count <- figure_count + 1

figure_number <- diabetes_agesex %>% 
  f_bar_grouped_agesex(y_variable = aboriginal_n,
                       y_breaks   = 10,
                       y_expand   = 7)

figure_percent <- diabetes_agesex %>% 
  f_bar_grouped_agesex(y_variable = aboriginal_prop,
                       y_max      = 50,
                       y_breaks   = 5,
                       y_expand   = 1)

f_bar_grouped_combine(figure_number, figure_percent)
```

#### Figure 10F-4

::: {#fig_10F_4 .fig-tab-title}
**Figure 10F-4:** Observed vs. expected number of Aboriginal and/or Torres Strait Islander people who have been told by a doctor or nurse that they have diabetes, NEPHU catchment, by age and sex, 2021
:::

```{r}
#
figure_count <- figure_count + 1

diabetes_agesex %>% 
  f_bar_obsexp_agesex(y_breaks = 10,
                      y_expand = 8)
```

#### Table 10F-2

::: {#tab_10F_2 .fig-tab-title}
**Table 10F-2:** Observed vs. expected number of Aboriginal and/or Torres Strait Islander people who have been told by a doctor or nurse that they have diabetes, NEPHU catchment, by age and sex, 2021
:::

```{r}
#
table_count  <- table_count + 1
table_number <- paste0("Table_", table_section, "_", sprintf("%.02d", table_count))

table <- diabetes_agesex %>%
  f_preptable_obsexp_agesex()

if (!file.exists(paste0(here::here(), "/data/", table_number, ".csv"))) {
  
  write.csv(table, paste0(here::here(), "/data/", table_number, ".csv"), row.names = FALSE)
  
}

table %>% 
  dplyr::filter(sex == "Female") %>%
  #
  dplyr::select(-sex) %>% 
  #
  f_table_obsexp(total_row   = nrow(table) / 2,
                 column_name = "Age group",
                 header_name = "Female")

table %>% 
  dplyr::filter(sex == "Male") %>%
  #
  dplyr::select(-sex) %>% 
  #
  f_table_obsexp(total_row   = nrow(table) / 2,
                 column_name = "Age group",
                 header_name = "Male")
```

<div style="height: 1em;"></div>

<a href="../data/`r paste0(table_number, '.csv')`" download class="btn btn-primary download-link">Download table data as CSV</a>

::::

::: {.callout-note title="Data notes" icon=false collapse=false}
`r note_data_census`

<div style="height: 1em;"></div>

`r note_caveat_underidentification`

`r note_caveat_unknown_agesex`

- Census respondents may have reported Type 1 and/or Type 2 diabetes in response to this question.
:::

### Hospital Admissions

```{r}
# ------------------------------------------------------------------------------
# Hospitalisations - diabetes
# Data from PHIDU Social Health Atlas
# ------------------------------------------------------------------------------
# Read in PHIDU data
hospital_diabetes <- f_read_phidu_iare_hospital(data_sheet    = "Admissions_principal_diag",
                                                column_select = "admissions_for_diabetes_aboriginal_people")

# Data wrangling
hospital_diabetes <- hospital_diabetes %>% 
  dplyr::mutate(iare_name = factor(iare_name, levels = iare_inc_vic_lvl)) %>% 
  #
  dplyr::arrange(iare_name)
```

- Data for this topic were obtained from the [PHIDU Social Health Atlas](https://phidu.torrens.edu.au/social-health-atlases).

- During the four-year period 01 July 2017 to 30 June 2021 there were 504 hospital admissions for diabetes for Aboriginal and/or Torres Strait Islander people living in Greater Melbourne.

- The number of admissions for diabetes for Aboriginal and/or Torres Strait Islander people living in different IAREs within the NEPHU catchment ranged from 118 (Northcote/Preston/Whittlesea IARE) to 5 (Knox and Melbourne East IAREs).

- The age-standardised rate of admissions for diabetes for Aboriginal and/or Torres Strait Islander people living in different IAREs within the NEPHU catchment ranged from 1,234 per 100,000 population (Maroondah IARE) to 66 per 100,000 population (Melbourne East IARE).

- 4 of the 10 IAREs in the NEPHU catchment area had an age-standardised admission rate for diabetes above the average rate for all Aboriginal and/or Torres Strait Islander people living in Greater Melbourne (Maroondah, Whitehorse, Northcote/Preston/Whittlesea, and Moreland/Broadmeadows IAREs).

::: {.callout-tip title="ICD-10-AM codes and definitions for this topic" icon=false collapse=false}
Admissions for diabetes are defined as those with a principal diagnosis code in the following list, coded using the [International Statistical Classification of Diseases and Related Health Problems 10th Revision Australian Modification (ICD-10-AM)](https://www.ihacpa.gov.au/health-care/classification/icd-10-amachiacs):

  - E10: Type 1 diabetes mellitus
  - E11: Type 2 diabetes mellitus
  - E12: Malnutrition-related diabetes mellitus
  - E13: Other specified diabetes mellitus
  - E14: Unspecified diabetes mellitus

<div style="height: 2em;"></div>
:::

:::: {.panel-tabset}

#### Figure 10F-5

::: {#fig_10F_5 .fig-tab-title}
**Figure 10F-5:** Number and age-standardised rate (per 100,000 population) of hospital admissions for diabetes for people who identified as Aboriginal and/or Torres Strait Islander, NEPHU catchment, by IARE, 01 July 2017 to 30 June 2021
:::

```{r}
#
figure_count <- figure_count + 1

figure_number <- hospital_diabetes %>% 
  f_bar_simple_iare(y_variable = aboriginal_n,
                    y_breaks   = 20,
                    y_expand   = 5,
                    hospital   = "admission")

figure_percent <- hospital_diabetes %>% 
  f_bar_simple_iare(y_variable = aboriginal_rate,
                    y_breaks   = 200,
                    y_expand   = 200,
                    hospital   = "admission")

f_bar_simple_combine(figure_number, figure_percent)
```

#### Table 10F-3

::: {#tab_10F_3 .fig-tab-title}
**Table 10F-3:** Hospital admissions for diabetes for people who identified as Aboriginal and/or Torres Strait Islander, NEPHU catchment, by IARE, 01 July 2017 to 30 June 2021
:::

```{r}
#
table_count  <- table_count + 1
table_number <- paste0("Table_", table_section, "_", sprintf("%.02d", table_count))

table <- hospital_diabetes %>%
  dplyr::select(-iare_name_short)

if (!file.exists(paste0(here::here(), "/data/", table_number, ".csv"))) {
  
  write.csv(table, paste0(here::here(), "/data/", table_number, ".csv"), row.names = FALSE)
  
}

table %>% 
  dplyr::mutate(aboriginal_rate = round(aboriginal_rate, digits = 0),
                hospital_ratio  = round(hospital_ratio, digits = 0)) %>% 
  #
  f_table_simple(iare     = "yes",
                 hospital = "admission")
```

<div style="height: 1em;"></div>

<a href="../data/`r paste0(table_number, '.csv')`" download class="btn btn-primary download-link">Download table data as CSV</a>

::::

::: {.callout-note title="Data notes" icon=false collapse=false}
`r note_data_phidu_vaed`

<div style="height: 1em;"></div>

`r note_caveat_underidentification`

`r note_caveat_iare_nephu`

<div style="height: 0.1em;"></div>

`r note_caveat_iare_ratio`

<div style="height: 1em;"></div>

`r note_caveat_admissions`

<div style="height: 2em;"></div>
:::

