---
title: "Infectious Diseases"
---

```{r}
#
source(here::here("setup.R"))
source(here::here("setup_footnotes.R"))

# Set up numbering for exporting tables as .csv
table_section <- "10K"
table_count   <- 0

figure_count <- 0
```

::: {.caveat-box}
**A note about the impact of under-identification**

`r note_box_underidentification`

**A note about notifiable diseases data**

`r note_box_notifiable`
:::

<div style="height: 2em;"></div>

::: {.key-question}
What infectious diseases affect Aboriginal and/or Torres Strait Islander people living in the NEPHU catchment?
:::

```{r}
#-------------------------------------------------------------------------------
# Notifiable conditions
# Data from PHESS
#-------------------------------------------------------------------------------
# Read in PHESS data
cases_all <- read.csv(paste0(here::here(), "/data", "/PHAR_Extract_2015_2024.csv")) 

# Data wrangling
cases_all <- cases_all %>% 
  dplyr::mutate(
    event_date = lubridate::ymd(event_date),
    year       = as.character(lubridate::year(event_date)),
    #
    lga = stringr::str_extract(lga, "^.*?(?=\\()"),
    lga = stringr::str_trim(lga),
    #
    nephu_case = dplyr::case_when(lga %in% nephu_lga ~ "NEPHU",
                                  TRUE ~ "Rest of Victoria"),
    #
    indigenous_status = dplyr::case_when(
      indigenous_status == "Aboriginal and Torres Strait Islander origin"     ~ "aboriginal",
      indigenous_status == "Aboriginal but not Torres Strait Islander origin" ~ "aboriginal",
      indigenous_status == "Torres Strait Islander but not Aboriginal origin" ~ "aboriginal",
      indigenous_status == "Not Aboriginal or Torres Strait Islander"         ~ "non_aboriginal",
    TRUE ~ "not_stated"),
    #
    condition = dplyr::case_when(
      condition == "Meningococcal infection"                  ~ "IMD",
      condition == "Pneumococcal infection (IPD)"             ~ "IPD",
      condition == "Varicella zoster infection (Chickenpox)"  ~ "VZV",
      condition == "Varicella zoster infection (Shingles)"    ~ "VZV",
      condition == "Varicella zoster infection (Unspecified)" ~ "VZV",
      #
      condition == "Gonococcal infection"            ~ "Gonococcal",
      condition == "Hepatitis B - Newly acquired"    ~ "HBV",
      condition == "Hepatitis B - Unspecified"       ~ "HBV",
      condition == "Hepatitis C - <24 months of age" ~ "HCV",
      condition == "Hepatitis C - Newly acquired"    ~ "HCV",
      condition == "Hepatitis C - Unspecified"       ~ "HCV",
      #
      condition == "Human Immunodeficiency Virus Infection - Child aged less than 18 months"     ~ "HIV",
      condition == "Human Immunodeficiency Virus Infection - Individual aged 18 months or older" ~ "HIV",
      #
      condition == "Acute Rheumatic Fever"          ~ "ARF",
      condition == "Invasive Group A Streptococcus" ~ "iGAS",
      condition == "Rheumatic Heart Disease"        ~ "RHD",
      TRUE ~ condition)) %>% 
  #
  dplyr::filter(indigenous_status != "not_stated")
```

## Vaccine Preventable Conditions

### Pertussis (Whooping Cough)

```{r}
#-------------------------------------------------------------------------------
# Notifiable conditions - pertussis
# Data from PHESS
#-------------------------------------------------------------------------------
pertussis <- cases_all %>% 
  f_wrangle_phess(condition_name = "Pertussis")
```

- Data for this topic were obtained from the [Victorian Public Health Event Surveillance System](https://www.health.vic.gov.au/infectious-diseases/infectious-diseases-surveillance-in-victoria).

- The number and rate of pertussis (whooping cough) notifications decreased during the COVID-19 pandemic and rebounded significantly during 2024. 

- This increase in 2024 was seen for both Aboriginal and/or Torres Strait Islander people and non-Indigenous people in both the NEPHU catchment and Victoria, and similar increases were seen across all other Australian states and territories.

::: {.callout-tip title="Notification and follow-up requirements" icon=false collapse=false}
- Pertussis (whooping cough) is a routine notifiable condition. Cases must be notified by medical practitioners and pathology services in writing to the Department of Health within five days of diagnosis.

<div style="height: 1em;"></div>

- The following pertussis notifications are signed out for case management and follow-up by Local Public Health Units:
  -	Children aged <12 months
  -	Women known to be in the last month of pregnancy
  -	Persons known to be in contact with children aged <6 months or women in the last month of pregnancy
  -	Persons known to attend or work in settings where there is likely to be contact with children aged <6 months or women in the last month of pregnancy
  - Depending on operational capacity, children aged <5 years and children aged <10 years should also be followed up

<div style="height: 1em;"></div>

- The management of pertussis transitioned from the Victorian Department of Health to Local Public Health Units on 18/07/2022.

<div style="height: 2em;"></div>
:::

:::: {.panel-tabset}

#### Figure 10K-1

::: {#fig_10K_1 .fig-tab-title}
**Figure 10K-1:** Number of pertussis notifications for Aboriginal and/or Torres Strait Islander people, Victoria, by year
:::

```{r}
#
figure_count <- figure_count + 1

pertussis %>% 
  f_bar_stacked_phess(y_max    = NA,
                      y_breaks = 20,
                      y_expand = 20)
```

#### Figure 10K-2

::: {#fig_10K_2 .fig-tab-title}
**Figure 10K-2:** Pertussis notification rate, Victoria, by Aboriginal and/or Torres Strait Islander status and year
:::

```{r}
#
figure_count <- figure_count + 1

pertussis %>%
  dplyr::rename(rate = vic_rate) %>% 
  #
  f_line_grouped_indigenous(y_variable = rate,
                            y_max      = NA,
                            y_breaks   = 50,
                            y_expand   = 60)
```

#### Table 10K-1

::: {#tab_10K_1 .fig-tab-title}
**Table 10K-1:** Number of pertussis notifications and notification rate, NEPHU and Victoria, by Aboriginal and/or Torres Strait Islander status and year
:::

```{r}
#
table_count  <- table_count + 1
table_number <- paste0("Table_", table_section, "_", sprintf("%.02d", table_count))

table <- pertussis %>% 
  dplyr::select(-indigenous_status,
                -notnephu_n) %>% 
  #
  tidyr::pivot_wider(names_from  = "indigenous_label",
                     values_from = c("nephu_n", "nephu_rate", "vic_n", "vic_rate"),
                     names_vary  = "slowest") %>%
  #
  janitor::clean_names()

if (!file.exists(paste0(here::here(), "/data/", table_number, ".csv"))) {
  
  write.csv(table, paste0(here::here(), "/data/", table_number, ".csv"), row.names = FALSE)
  
}

table %>% 
  dplyr::mutate(nephu_rate_aboriginal     = round(nephu_rate_aboriginal, digits = 1),
                vic_rate_aboriginal       = round(vic_rate_aboriginal, digits = 1),
                nephu_rate_non_indigenous = round(nephu_rate_non_indigenous, digits = 1),
                vic_rate_non_indigenous   = round(vic_rate_non_indigenous, digits = 1)) %>%
  #
  f_table_grouped_phess()
```

<div style="height: 1em;"></div>

<a href="../data/`r paste0(table_number, '.csv')`" download class="btn btn-primary download-link">Download table data as CSV</a>

::::

::: {.callout-note title="Data notes" icon=false collapse=false}
`r note_data_phess`

<div style="height: 1em;"></div>

`r note_caveat_underidentification`

`r note_caveat_notification_rate`

`r note_caveat_pandemic_decrease`
:::

### Invasive Meningococcal Disease (IMD)

```{r}
#-------------------------------------------------------------------------------
# Notifiable conditions - invasive meningococcal disease
# Data from PHESS
#-------------------------------------------------------------------------------
imd <- cases_all %>% 
  f_wrangle_phess(condition_name = "IMD")
```

- Data for this topic were obtained from the [Victorian Public Health Event Surveillance System](https://www.health.vic.gov.au/infectious-diseases/infectious-diseases-surveillance-in-victoria).

- The most recent notification of invasive meningococcal disease (IMD) for an Aboriginal and/or Torres Strait Islander resident of the NEPHU catchment was in 2018.

- For Victoria overall, the number of notifications and notification rates for IMD have decreased in the years following the COVID-19 pandemic, however the rates of IMD remain higher for Aboriginal and/or Torres Strait Islander people compared to non-Indigenous people.

::: {.callout-tip title="Notification and follow-up requirements" icon=false collapse=false}
- Invasive meningococcal disease (IMD) is an urgent notifiable condition. Cases must be immediately notified upon initial diagnosis (suspected or confirmed) by medical practitioners and pathology services by telephone to the Department of Health, and pathology services must follow-up with written notification within five days of diagnosis.

- All IMD notifications are signed out for urgent case management and follow-up by Local Public Health Units.

- The management of IMD transitioned from the Victorian Department of Health to Local Public Health Units on 27/02/2023.
:::

:::: {.panel-tabset}

#### Figure 10K-3

::: {#fig_10K_3 .fig-tab-title}
**Figure 10K-3:** Number of invasive meningococcal disease notifications for Aboriginal and/or Torres Strait Islander people, Victoria, by year
:::

```{r}
#
figure_count <- figure_count + 1

imd %>% 
  f_bar_stacked_phess(y_max    = 5,
                      y_breaks = 1,
                      y_expand = 0.25)
```

#### Figure 10K-4

::: {#fig_10K_4 .fig-tab-title}
**Figure 10K-4:** Invasive meningococcal disease notification rate, Victoria, by Aboriginal and/or Torres Strait Islander status and year
:::

```{r}
#
figure_count <- figure_count + 1

imd %>%
  dplyr::rename(rate = vic_rate) %>% 
  #
  f_line_grouped_indigenous(y_variable = rate,
                            y_max      = NA,
                            y_breaks   = 1,
                            y_expand   = 0.65)
```

#### Table 10K-2

::: {#tab_10K_2 .fig-tab-title}
**Table 10K-2:** Number of invasive meningococcal disease notifications and notification rate, NEPHU and Victoria, by Aboriginal and/or Torres Strait Islander status and year
:::

```{r}
#
table_count  <- table_count + 1
table_number <- paste0("Table_", table_section, "_", sprintf("%.02d", table_count))

table <- imd %>% 
  dplyr::select(-indigenous_status,
                -notnephu_n) %>% 
  #
  tidyr::pivot_wider(names_from  = "indigenous_label",
                     values_from = c("nephu_n", "nephu_rate", "vic_n", "vic_rate"),
                     names_vary  = "slowest") %>%
  #
  janitor::clean_names()

if (!file.exists(paste0(here::here(), "/data/", table_number, ".csv"))) {
  
  write.csv(table, paste0(here::here(), "/data/", table_number, ".csv"), row.names = FALSE)
  
}

table %>% 
  dplyr::mutate(nephu_rate_aboriginal     = round(nephu_rate_aboriginal, digits = 2),
                vic_rate_aboriginal       = round(vic_rate_aboriginal, digits = 2),
                nephu_rate_non_indigenous = round(nephu_rate_non_indigenous, digits = 2),
                vic_rate_non_indigenous   = round(vic_rate_non_indigenous, digits = 2)) %>%
  #
  f_table_grouped_phess()
```

<div style="height: 1em;"></div>

<a href="../data/`r paste0(table_number, '.csv')`" download class="btn btn-primary download-link">Download table data as CSV</a>

::::

::: {.callout-note title="Data notes" icon=false collapse=false}
`r note_data_phess`

<div style="height: 1em;"></div>

`r note_caveat_underidentification`

`r note_caveat_notification_rate`

`r note_caveat_pandemic_decrease`
:::

