---
title: "Mental Health and Wellbeing"
---

```{r}
#
source(here::here("setup.R"))
source(here::here("setup_footnotes.R"))

# Set up numbering for exporting tables as .csv
table_section <- "10L"
table_count   <- 0

figure_count <- 0
```

::: {.caveat-box}
**A note about the impact of under-identification**

`r note_box_underidentification`

**A note about interpreting observed vs. expected figures**

`r note_box_obsexp`
:::

<div style="height: 2em;"></div>

::: {.key-question}
How do Aboriginal and/or Torres Strait Islander people living in the NEPHU catchment experience emotional wellbeing?
:::

## Cultural Strengths and Emotional Wellbeing

### Positive Wellbeing

```{r}
# ------------------------------------------------------------------------------
# Positive wellbeing score
# Data from ABS National Aboriginal and Torres Strait Islander Health Survey
# ------------------------------------------------------------------------------
# Read in NATSIHS data
wellbeing_positive <- f_read_natsihs_sex("Indicator_Vic_Sex_Resilience_PositiveWellbeing_Score.csv")

# Data wrangling
wellbeing_positive <- wellbeing_positive %>% 
  dplyr::rowwise() %>% 
  dplyr::mutate(answered_n = sum(across(high_75_100_n:low_0_25_n))) %>% 
  dplyr::ungroup() %>% 
  #
  janitor::adorn_totals(fill = "Total") %>% 
  #
  dplyr::rename(high_n   = high_75_100_n,
                medium_n = medium_26_74_n,
                low_n    = low_0_25_n) %>% 
  #
  dplyr::mutate(across(high_n:low_n,
                       ~ . / answered_n * 100,
                       .names = "{.col}_prop"),
                #
                answered_prop = answered_n / answered_n * 100) %>% 
  #
  dplyr::rename_with(~ stringr::str_replace(., "_n_prop$", "_prop"), ends_with("_n_prop")) %>% 
  #
  dplyr::select(sex,
                starts_with("high"),
                starts_with("medium"),
                starts_with("low"),
                starts_with("answered")) %>% 
  #
  tidyr::pivot_longer(!sex,
                      names_to = c("answer", ".value"),
                      names_pattern = "(.+)_(.+)") %>% 
  #
  dplyr::mutate(answer = factor(dplyr::recode(answer,
                                              "high"     = "High",
                                              "medium"   = "Medium",
                                              "low"      = "Low",
                                              "answered" = "Total"),
                                levels = c("High",
                                           "Medium",
                                           "Low",
                                           "Total")))
```

- Data for this topic were obtained from the [2022-23 ABS National Aboriginal and Torres Strait Islander Health Survey](https://www.abs.gov.au/statistics/people/aboriginal-and-torres-strait-islander-peoples/national-aboriginal-and-torres-strait-islander-health-survey/latest-release).

- Overall, 30% of Aboriginal and/or Torres Strait Islander survey respondents aged 15 years and over living in Greater Melbourne had a high level of positive wellbeing in the four weeks prior to the survey. This equates to approximately 7,700 Aboriginal and/or Torres Strait Islander people.

- Females were more likely to have a low level of positive wellbeing than males (12% vs. 5.1%).

::: {.callout-tip title="Question/s asked for this topic" icon=false collapse=false}
The following questions were asked of all people aged 15 years and over completing the 2022-23 ABS National Aboriginal and Torres Strait Islander Health Survey. The responses to these four questions were combined and scaled to produce an overall score between 0 and 100. Overall scores were classified as Low (0 to 25), Medium (26 to 74), and High (75 to 100) positive wellbeing.

- In the last 4 weeks, about how often did you feel calm and peaceful?
  - All of the time
  - Most of the time
  - Some of the time
  - A little of the time
  - None of the time
  - Don't know
  - Prefer not to answer

<div style="height: 1em;"></div>

- In the last 4 weeks, about how often have you been a happy person?
  - All of the time
  - Most of the time
  - Some of the time
  - A little of the time
  - None of the time
  - Don't know
  - Prefer not to answer
  
<div style="height: 1em;"></div>

- In the last 4 weeks, about how often did you feel full of life?
  - All of the time
  - Most of the time
  - Some of the time
  - A little of the time
  - None of the time
  - Don't know
  - Prefer not to answer
  
<div style="height: 1em;"></div>

- In the last 4 weeks, about how often did you have a lot of energy?
  - All of the time
  - Most of the time
  - Some of the time
  - A little of the time
  - None of the time
  - Don't know
  - Prefer not to answer

<div style="height: 2em;"></div>
:::

:::: {.panel-tabset}

#### Figure 10L-1

::: {#fig_10L_1 .fig-tab-title}
**Figure 10L-1:** Level of positive wellbeing in the four weeks prior to the survey for people aged 15 years and over who identified as Aboriginal and/or Torres Strait Islander, Greater Melbourne, by sex, 2022-2023
:::

```{r}
#
figure_count <- figure_count + 1

figure_number <- wellbeing_positive %>% 
  dplyr::filter(sex != "Total" & answer != "Total") %>%
  #
  f_bar_grouped_sex(x_variable = answer,
                    y_variable = n,
                    y_breaks   = 1000,
                    y_expand   = 400)

figure_percent <- wellbeing_positive %>% 
  dplyr::filter(sex != "Total" & answer != "Total") %>%
  #
  f_bar_grouped_sex(x_variable = answer,
                    y_variable = prop,
                    y_max      = 100,
                    y_breaks   = 10,
                    y_expand   = 2.5)

f_bar_grouped_combine(figure_number, figure_percent, legend_offset = -0.1)
```

#### Table 10L-1

::: {#tab_10L_1 .fig-tab-title}
**Table 10L-1:** Level of positive wellbeing in the four weeks prior to the survey for people aged 15 years and over who identified as Aboriginal and/or Torres Strait Islander, Greater Melbourne, by sex, 2022-2023
:::

```{r}
#
table_count  <- table_count + 1
table_number <- paste0("Table_", table_section, "_", sprintf("%.02d", table_count))

table <- wellbeing_positive %>% 
  tidyr::pivot_wider(names_from  = "sex",
                     values_from = c("n", "prop"),
                     names_vary  = "slowest") %>% 
  #
  janitor::clean_names() %>% 
  #
  dplyr::rename(positive_wellbeing_score = answer)

if (!file.exists(paste0(here::here(), "/data/", table_number, ".csv"))) {
  
  write.csv(table, paste0(here::here(), "/data/", table_number, ".csv"), row.names = FALSE)
  
}

table %>% 
  dplyr::mutate(prop_male   = sprintf("%.1f", prop_male),
                prop_female = sprintf("%.1f", prop_female),
                prop_total  = sprintf("%.1f", prop_total)) %>%
  #
  f_table_grouped_sex(n_variable  = "estimate",
                      column_name = "Positive wellbeing")
```

<div style="height: 1em;"></div>

<a href="../data/`r paste0(table_number, '.csv')`" download class="btn btn-primary download-link">Download table data as CSV</a>

::::

::: {.callout-note title="Data notes" icon=false collapse=false}
`r note_data_natsihs`

<div style="height: 1em;"></div>

`r note_caveat_underidentification`

`r note_caveat_survey_unknown`

`r note_caveat_survey_estimate`
:::

